Likelihood ratio tests constructed with discriminative classifiers and calibrated with generative models

### 2014-11-20 tex/parametrized-learning.tex
references

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb, amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

@@

@@

>>>>

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb, amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{xfrac}

@@

\section{Comments on classification and frequentist hypothesis tests}

Significant literature exists around generative classifiers and discriminative classifiers. Typically, generative classifiers learn a model for the joint probability $p(x,y)$, of the inputs $x$ and the classification label $y$, and predict $p(y|x)$ via Bayes rule. In contrast, discriminative classifiers model the posterior $p(y|x)$ directly. For classification tasks, one then thresholds on $p(y|x)$. In both cases this description in terms of a posterior requires a prior distribution for $p(y)$, which is either modeled explicitly or learned from the training data. 
This formulation of classification is somewhat confusing in the setting of the current work.

First of all, the generative model $f(x|\theta)$ here is a generative statistical model, not a generative classifier. We think of the  $f(x|\theta)$ along the lines of a traditional scientific theory, able to make predictions about $x$ and being motivated by domain-specific considerations. For example, in the context of high energy particle physics $f(x|\theta)$ is based on quantum field theory and a detailed simulation of the particle detector and data processing algorithms that transform raw sensor data into the feature vector $x$. The point here is to make the distinction of $f(x|\theta)$ in this context from a generic generative classifier like normal discriminate analysis or naive Bayes.

Secondly, we are not interested in thresholded classification on individual events $x_e$, but are interested in a thresholded hypothesis test on an entire data set based on the likelihood ratio test statistic $T(D)$.  
Additionally, we know that both discriminative and generative classifier scores are often poorly calibrated. 
We wish to have well calibrated statements, but not about $p(y|x)$ but p-values associated to $P(T(D) > k |\theta)$.

Lastly, in the setting of frequentist hypothesis we do not have a prior $\pi(\theta)$. 
While we can use the generative models to produce training data $(x,y=0)$ generated 
from $f(x|\theta_0)$ and $(x,y=1)$ generated from $f(x|\theta_1)$, the relative mix $p(y)$ 
is arbitrary.  When $p(y=0)=p(y=1)=\sfrac{1}{2}$, then 
\begin{equation}
p(y=1 | x) = \frac{p(x|y=1)}{p(x|y=0)+p(x|y=1)} = \frac{f(x|\theta_1)}{f(x|\theta_0)+f(x|\theta_1))} \;,
\end{equation}
which is monotone with the desired likelihood ratio $f(x|\theta_1)/f(x|\theta_0)$ and free from the
complications of the prior $\pi(\theta)$ that is undefined in the frequentist context.

Since the prior $p(y)$ has no clear meaning for the test at hand and because the classifier score $p(x|y)$ may not be well calibrated, we choose to denote the classifier score $s(x)$ and simply think of it as a deterministic dimensionality reduction map $s: X \to \mathbb{R}$.

@@

Related work:
"Classification with hybrid generative/discriminative models"

Refs to include:
\cite{ClaytonScott,AndrewY.Ng,SamT.Roweis,EricP.Xing,SujaySanghavi,BiancaZadrozny,Zadrozny2001,TommiJaakkola}

@@

\bibliographystyle{unsrt}
\bibliography{learning.bib}
%\begin{thebibliography}
%\parametrized-learning}
%\end{thebibliography
